Selective electroporation of the viral envelope via nanosecond pulses

### 2021-03-21 documents/EPA_paper.tex
Writing notes
Pure electroporation

There is an interesting citation in the Gray literature \cite{Indexa}, particularly because it seems to have progressed to the clinic:

@@

\end{document}

>>>>

\section{Gray literature}

There is an interesting citation in the Gray literature \cite{Index1981}, particularly because it seems to have progressed to the clinic:

@@

Thanks to [] at york, and [] on the UWaterloo Discord.
\end{document}

### 2021-03-21 documents/summary.tex
Pure electroporation

However, long pulses at $\approx10$ MV/m appear to electroporate large enveloped viruses, similarly to host cells\cite{AC2017}, which we believe might hint that a regime of intensity, duration, and pulse count may exist to target the viral envelope. In fact, despite considering the somewhat harder problem of electroporating while inside the cell, rather than in extracellular fluid, \cite{Electroporation2013} report selective electroporation of 250 nm liposomes with 4 ns pulses up to 10 MV/m without inducing damage to the host cell - which seems a very important result. Despite the number of free parameters in pulse exposure, it is possible that scaling laws might conceivably preclude the formation of destructive transmembrane potentials in 100 nm-scale viruses.

@@

However, perhaps regrettably, the experiments reported in this study were not designed to address the question of whether conventional ion- and bilayer- driven electroporation can be harnessed, but rather to test whether purely mechanical damage from $qE$ forces can be easily observed in viruses.

@@

In-vitro, there appears to be a consistent set of effects of nanosecond pulses, or nsPEF, distinct from membrane effects, that occur on small intracellular organelles (see reviews \cite{Penetration2016c} \cite{Effects2016} ), \cite{Bioeffects},

@@

A rotational drag coefficient can extracted from MD simulation\cite{Evaluating2017}. We simulated a 30 nm lipid liposome using the MARTINI coarse-grained \cite{MARTINI2007}\cite{BUMPy2018}\cite{Tsjerk2020} lipid force fields with GROMACS\cite{GROMACS2005}' torque measurement feature, kindly implemented by Kutzner \cite{Keep2011}. This suggests a maximum rotation frequency on the order of $10^3$, but this value requires so many assumptions that it's hardly worth mentioning. From \cite{Spinning2008}, the upper bound of a homogenous spherical virus with 10 MPa overall tensile strength appears to be about $10^9$ Hz. \footnote{\ghfile{documents/optical_centrifuge.ipynb}}

@@

If the effect exists, efficient use of the lower damping would seem to require more than one cycle of oscillation. It appears to be impossible at any input power to effectively drive a 10 GHz resonator in the lungs from outside the torso, partly because of the layer of conductive muscle that shields the chest cavity (see supplemental). Open-ended waveguides\cite{OpenEnded1982}\cite{Analysis1989}\cite{142018} or water bolus allow efficient coupling to the body. There is precedent for bronchoscopic \cite{Flexible2019}\cite{Antenna2018} or more invasive methods\cite{Implantable1980}\cite{Implantable1982}\cite{Electromagnetic1983}.

@@

@@

Fecher\cite{Production1949} micro-oscillographs

@@

A clone of Li et al's pulse generator was engraved with a 30 degree V-bit cutter onto a 1.6 mm double-sided 1 oz copper FR4 substrate. ATC 100B470JT500XT capacitors and FMMT417 transistors were used.

@@

HIA4 power supply.

>>>>

However, long pulses at $\approx10$ MV/m appear to electroporate large enveloped viruses, similarly to host cells\cite{AC2017}, which we believe might hint that a regime of intensity, duration, and pulse count may exist to target the viral envelope. In fact, despite considering the somewhat harder problem of electroporating while inside the cell, rather than in extracellular fluid, \cite{Electroporation2013} report that 250 nm liposomes should be porated without inducing damage to the host cell, and modelling shows that 100 nm liposomes may be porated by 4 ns pulses never exceeding 20 MV/m. Despite the number of free parameters in pulse exposure, it is possible that scaling laws might conceivably preclude the safe formation of destructive transmembrane potentials in 100 nm-scale viruses.

@@

However, perhaps regrettably, the experiments reported in this study were not designed to address the question of whether this conventional ion- and bilayer- driven electroporation can be harnessed, but rather to test whether purely mechanical damage from $qE$ forces can be easily observed in viruses.

@@

In-vitro, there appears to be a consistent set of effects of nanosecond pulses, or nsPEF, distinct from membrane effects, that occur on small intracellular organelles (see reviews \cite{Penetration2016c} \cite{Effects2016} ), \cite{Bioeffects}.

Apoptosis \cite{Effect2008}

@@

A rotational drag coefficient can extracted from MD simulation\cite{Evaluating2017}. We simulated a 30 nm lipid liposome using the MARTINI coarse-grained \cite{MARTINI2007}\cite{BUMPy2018}\cite{Tsjerk2020} lipid force fields with GROMACS\cite{GROMACS2005}' torque measurement feature, kindly implemented by Kutzner \cite{Keep2011}. This suggests a maximum rotation frequency on the order of $10^3$, but this value requires so many assumptions that it's hardly worth mentioning.

From \cite{Spinning2008}, the upper bound of a homogenous spherical virus with 10 MPa overall tensile strength appears to be about $10^9$ Hz. \footnote{\ghfile{documents/optical_centrifuge.ipynb}}

@@

If the effect exists, efficient use of the lower damping would seem to require more than one cycle of oscillation. It appears to be impossible at any input power to effectively drive a 10 GHz resonator in the lungs from outside the torso, partly because of the layer of conductive muscle that shields the chest cavity.

Open-ended waveguides\cite{OpenEnded1982}\cite{Analysis1989}\cite{142018} or water bolus allow efficient coupling to the body. There is precedent for bronchoscopic \cite{Flexible2019}\cite{Antenna2018} or more invasive methods\cite{Implantable1980}\cite{Implantable1982}\cite{Electromagnetic1983}.

@@

See supplemental for a cursory treatment of the Brillouin precursors.

@@

%Fecher\cite{Production1949} micro-oscillographs

@@

A clone of Li et al's pulse generator was engraved with a 30 degree V-bit cutter onto a 1.6 mm double-sided 1 oz copper FR4 substrate. ATC 100B470JT500XT capacitors and FMMT417 transistors were used. This was powered by a HIA4 power supply at 330v.

@@
